% ************************** Thesis Acknowledgements **************************

\begin{acknowledgements}      

Throughout the process of writing this thesis I had the opportunity to reflect on these last 4 years and how thankful and privileged I am for having the people around me that made this all possible. First, I would like to thank the John Innes Centre and the NRPDTP programme for providing the perfect environment to carry out this research. I felt so spoilt by all the support and guidance offered by Lab Support, Bioimaging, Horticultural Sevices and Research Computing, thank you all!

I want to thank Christine Faulkner for offering insightful, impartial feedback on my PhD, whilst also being encouraging, realistic and empathetic. To my supervisor Xiaoqi Feng - thank you for always sharing my excitement for my project, allowing me to take the reign in exploring topics that excited me and for your unwavering faith in my abilities, both personal and scientific. 

I am also deeply grateful to the members of the Feng lab, past and present from whom I've have learnt so much. To the lovely postdocs Yalin and Shaoli, it was so much fun to work with you and your work ethic and enthusiasm never fails to impress me. A big thank you to Shujuan an Chris, who always looked out for me after the lab moved abroad, your presence in the lab made all the difference. You were both always so approachable, easygoing and were always up for helping me or just having a much needed pep talk or chat. I owe a special thanks to Jincheng Long who taught me everything I know about sRNAs and techniques in the lab - you were an amazing mentor in my first two years and so much fun to be around!  I also want to thank Martin Vickers for making bioinformatics approachable and enjoyable — it became the highlight of my PhD thanks to you.

To my adoptive lab, the Carella Lab -  Hetty, Hyeonmin, Bea, Ellie, Sam and Kayla - you are all amazing! You made me feel so welcome in the lab and I will miss our lunches and silly (and serious) conversations. I owe a massive thanks to Phil Carella - thank you so much for agreeing to be my supervisor for the final part of my PhD. Your kindness, support, and calm presence made all the difference.

I want to thank my friends, Sophie, Deirdre, Cat, Joe and especially Emily, you have been there through the ups and downs of these four years, I feel so privileged to have found friends like you. Your kindness, humour, enthusiasm and infectious passion for science has made living in Norwich around you amazing. I will miss our dinner parties, coffees and hangouts so much. I’m so proud of all of you for reaching the end of your PhDs!

Köszönöm a családomnak: Papa, Mama, Andris, Zsófi, Zsuzsi, Bogesz, Dorka, Balu, Zsuzsa, Nagyi és Nagypapa köszönöm a sok segítséget, bíztatást és támogatást, nélkületek nem sikerült volna! Nagyon szeretlek titeket!

To Johnny, thank you for being my cheerleader, you have been the most important rock throughout these 4 years. Thank you for believing in me and for moving to Norwich so I could continue to pursue this degree. Thank you for cooking for me, and offering distractions when I needed them most.

Lastly, I want to give a huge thank you to Jimmy. You have been the most amazing, patient and invaluable mentor and friend throughout my PhD. The amount of dedication, kindness and enthusiasm you showed towards my project is absolutely incredible, even after you moved away to the other side of the pond. I am so grateful for your guidance and friendship. You made it possible for me to see this project through, and I truly couldn’t have completed this degree without you.

Thank you all!


\end{acknowledgements}
